Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-04-14 ltx/BrainIDE.tex
UKF algorithm

The approximation of the state density is made by propagating the so called sigma points through the state-space model. The sigma points describe the mean and covariance of the state estimate. The accuracy of the approximation is linked to the time step used in the temporal discretization of the system, where the state transition appears more linear with a finer resolution. To explain the operation of the UKF we begin by deriving the traditional Kalman filter with the state-space model
\begin{eqnarray}
	\mathbf{x}_t = q(\mathbf{x}_t)\boldsymbol{\theta} - \lambda\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{eqnarray}
where $\boldsymbol{\theta}$ and $\lambda$ are the systems parameters. In forwards iteration since the observation equation is linear in parameters we only need to use Unsented transformation in prediction step. The update step equations remains the same as the standard Kalman filtering formulations. In backward smoothing path, estimation from the forward filtering is used to form the new set of sigma points in a similar way to the filterin step, however the predicted cross-covariance of the states need to be calculated. A summary of the algorithm is given in Algorithm 
\begin{algorithm}
\caption{The Unscented Kalman Filter}\label{CPTEMDAlgorithm}
\begin{algorithmic}[1]
\State Initialise
\begin{equation*}
\hat{\mathbf x}_0=\mathbb E[\mathbf x_0]
\end{equation*}
\begin{equation*}
\mathbf P_0=\mathbb E[(\mathbf x_0-\hat{\mathbf x}_0)(\mathbf x_0-\hat{\mathbf x}_0)^\top]
\end{equation*}
\State Form the sigma vectors 
\begin{equation*}
\mathcal X_k=[\hat{\mathbf x}_k \quad \hat{\mathbf x}_k+\gamma\sqrt{\mathbf P_k} \quad \hat{\mathbf x}_k-\gamma\sqrt{\mathbf P_k}]
\end{equation*}
\State Update the sigma vectors by propogating them through state transition equation and compute the predicted state and the predicted state covariance
\begin{equation*}
 \hat{\mathbf x}_{k +1|k}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,k+1|k}
\end{equation*}
\begin{equation*}
 \mathbf P_{k +1|k}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,k+1|k}-\hat{\mathbf x}_{k +1|k})(\mathcal X_{i,k+1|k}-\hat{\mathbf x}_{k +1|k})^\top+\boldsymbol \Sigma_w
\end{equation*}
\State use extrema to form the splines $s_{min,k}(n)$ and $s_{max,k}(n)$
\State calculate $x_{0,k}(n) = 0.5\left(s_{min,k}(n) + s_{max,k}(n)\right)$
\State find local maxima of $h_k(n)$
\State use maxima to form the spline $r_k(n) = s_{max,k}(n)$
\While { $x_{0,k}(n)$ is not monotonic }
    \State $k=k+1$
    \State perform CPT on $x_{0,k-1}(n)$ to estimate $\phi_{k}(n)$
    \State repeats steps 3 to 7 to approximate parameters
\EndWhile

\end{algorithmic}
\end{algorithm}
\section{Comparison of Estimated Parameters and True Parameters}

>>>>

The approximation of the state density is made by propagating the so called sigma points through the state-space model. The sigma points describe the mean and covariance of the state estimate. The accuracy of the approximation is linked to the time step used in the temporal discretization of the system, where the state transition appears more linear with a finer resolution. To explain the operation of the UKF we begin by deriving the traditional Kalman filter with the state-space model
\begin{eqnarray}
	\mathbf{x}_t = q(\mathbf{x}_t)\boldsymbol{\theta} - \lambda\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{eqnarray}
where $\boldsymbol{\theta}$ and $\lambda$ are the systems parameters. In forwards iteration since the observation equation is linear in parameters we only need to use Unsented transformation in prediction step. The update step equations remains the same as the standard Kalman filtering formulations. In backward smoothing path, estimation from the forward filtering is used to form the new set of sigma points in a similar way to the filterin step, however the predicted cross-covariance of the states need to be calculated. A summary of the algorithm is given in Algorithm 
\begin{algorithm}
\caption{The Unscented Kalman Filter}\label{CPTEMDAlgorithm}
\begin{algorithmic}[1]
\State Initialise
\begin{equation*}
\hat{\mathbf x}_0, \mathbf P_0
\end{equation*}
\State Forward iteration
For $t \in \left\lbrace 0,\cdots, N\right\rbrace $,
\begin{equation*}
\mathcal X_t^f=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^{f}-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
Propogate the sigma points through the dynamic model
\begin{equation}
\mathcal X_{i,t+1|t}^{f*}=Q(\mathcal X_t^f,u_t)
\end{equation}
\begin{equation*}
 \hat{\mathbf x}_{k +t}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1|t}^{f*}
\end{equation*}
\begin{equation*}
 \mathbf P_{t +1}^-=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e
\end{equation*}
\begin{equation*}
\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_v)^{-1}
\end{equation*}
\begin{equation*}
\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-})
\end{equation*}
\begin{equation*}
 \mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}
\end{equation*}
\State Backward iteration
\State initialisation
\begin{equation}
 \mathbf P_T^s= \mathbf P_T^f
\end{equation}
\begin{equation}
\hat{\mathbf x}^s_T= \hat{\mathbf x}^f_T
\end{equation}
For $t \in \left\lbrace N-1, \cdots, 0 \right\rbrace $
\begin{equation*}
\mathcal X_t^s=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^f-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
\end{algorithmic}
\end{algorithm}
\section{Comparison of Estimated Parameters and True Parameters}
